---
title: "Bile predicaments"
author: "HCS"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    theme: united
    toc_float: yes
    toc: yes
    number_sections: yes
    code_folding: hide
  word_document:
    toc: yes
  pdf_document:
    toc: yes
  html_notebook:
    theme: united
    toc_float: yes
    toc: yes
    number_sections: yes
    code_folding: hide
---


```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```


# Prepare analysis
## Required packages
Load the R packages required for the analysis.
```{r}
library(tidyverse)     # data handling
library(openxlsx)      # read/write xlsx files
library(readxl)        # read xlsx files
library(compareGroups) # Descriptive statistics and tables
library(cardx)         # Analyses for R
library(gtsummary)     # Create tables
library(caret)         # Calculate sensitivity, specificity and predictive values 
library(ggpubr)        # Visualization
library(ggsci)
library(FactoMineR)    # Visualization
library(factoextra)    # Visualization
library(ggmosaic)      # Visualization
library(kableExtra)    # Render tables
library(ROCit)         # ROC curves
```



## Auxiliary functions
Custom functions used in the analysis.
```{r}
# Compute performance metrics of diagnostic tests
# Cmompute 95% CIs for all parameters

#' Compute performance metrics (sensitivity, specificity, PPV, NPV, accuracy, LRs, F1)
#' with 95% confidence intervals.
#'
#' @param confusion_mtx 2x2 matrix or table. Rows = predicted classes, columns = reference (actual) classes
#'                      (default common layout produced by caret::confusionMatrix when passing (pred, ref)).
#'                      If your layout differs, supply the mapping via the index arguments or set
#'                      argument layout = c("pred_rows_ref_cols", "ref_rows_pred_cols").
#' @param prevalence Optional numeric giving disease prevalence (for PPV/NPV if you want to override sample prevalence).
#' @param pos Name of the positive class.
#' @param neg Name of the negative class.
#' @param ci_level Confidence level (default 0.95).
#' @param ci_method One of "wilson", "wald", "exact".
#' @param layout Either "pred_rows_ref_cols" (default) or "ref_rows_pred_cols".
#' @param add_additional If TRUE add accuracy, likelihood ratios, F1 (with CIs where applicable).
#' @param clamp Whether to clamp CI limits to [0,1] for proportion metrics.
#' @return Named numeric vector (or list) of metrics with CIs.
#' @importFrom stats binom.test
#' @examples
#' # Example (assumes you have caret installed for sensitivity()/specificity()/posPredValue()/negPredValue())
#' # library(caret)
#' # m <- matrix(c(50,10,5,200), nrow=2,
#' #             dimnames=list(Predicted=c("yes","no"), Actual=c("yes","no")))
#' # get_perf(m, pos="yes", neg="no")
get_perf <- function(confusion_mtx,
                     prevalence = NULL,
                     pos = "yes",
                     neg = "no",
                     ci_level = 0.95,
                     ci_method = c("wilson","wald","exact"),
                     layout = c("pred_rows_ref_cols","ref_rows_pred_cols"),
                     add_additional = TRUE,
                     clamp = TRUE) {

  ci_method <- match.arg(ci_method)
  layout    <- match.arg(layout)

  # Basic validation
  if (!is.matrix(confusion_mtx) && !is.table(confusion_mtx)) {
    stop("confusion_mtx must be a 2x2 matrix or table.")
  }
  if (any(dim(confusion_mtx) != 2)) {
    stop("confusion_mtx must be 2x2.")
  }

  # Ensure dimnames
  dn <- dimnames(confusion_mtx)
  if (is.null(dn) || any(vapply(dn, is.null, logical(1)))) {
    stop("confusion_mtx must have dimnames for both dimensions (predicted and actual classes).")
  }

  # Reorient if needed
  # After this block we want rows = predicted, columns = actual
  if (layout == "ref_rows_pred_cols") {
    confusion_mtx <- t(confusion_mtx)
  }

  if (!(pos %in% rownames(confusion_mtx)) || !(pos %in% colnames(confusion_mtx))) {
    stop("Positive class 'pos' must appear in both row and column names.")
  }
  if (!(neg %in% rownames(confusion_mtx)) || !(neg %in% colnames(confusion_mtx))) {
    stop("Negative class 'neg' must appear in both row and column names.")
  }

  # Extract cells
  # Convention:
  # TP: predicted = pos, actual = pos
  # FP: predicted = pos, actual = neg
  # FN: predicted = neg, actual = pos
  # TN: predicted = neg, actual = neg
  TP <- confusion_mtx[pos, pos]
  FP <- confusion_mtx[pos, neg]
  FN <- confusion_mtx[neg, pos]
  TN <- confusion_mtx[neg, neg]

  N  <- TP + FP + FN + TN

  # Helper for proportion CIs
  binom_ci <- function(success, total, level, method) {
    if (total == 0) return(c(NA, NA))
    alpha <- 1 - level
    if (method == "wald") {
      p <- success / total
      z <- qnorm(1 - alpha/2)
      se <- sqrt(p*(1-p)/total)
      lower <- p - z*se
      upper <- p + z*se
    } else if (method == "wilson") {
      p <- success / total
      z <- qnorm(1 - alpha/2)
      denom <- 1 + z^2/total
      center <- p + z^2/(2*total)
      half <- z * sqrt(p*(1-p)/total + z^2/(4*total^2))
      lower <- (center - half)/denom
      upper <- (center + half)/denom
    } else if (method == "exact") {
      bt <- binom.test(success, total, conf.level = level)
      lower <- bt$conf.int[1]
      upper <- bt$conf.int[2]
    }
    if (clamp) {
      lower <- max(0, lower)
      upper <- min(1, upper)
    }
    c(lower, upper)
  }

  # Proportion + CI wrapper
  prop_with_ci <- function(success, total) {
    p <- if (total == 0) NA_real_ else success / total
    ci <- binom_ci(success, total, ci_level, ci_method)
    list(estimate = p, lower = ci[1], upper = ci[2])
  }

  # Core metrics
  sens <- prop_with_ci(TP, TP + FN)          # Sensitivity
  spec <- prop_with_ci(TN, TN + FP)          # Specificity
  ppv  <- prop_with_ci(TP, TP + FP)          # PPV
  npv  <- prop_with_ci(TN, TN + FN)          # NPV
  acc  <- prop_with_ci(TP + TN, N)           # Accuracy

  # Prevalence (sample vs provided) (used only for reporting)
  prev_sample <- (TP + FN)/N
  prev <- if (is.null(prevalence)) prev_sample else prevalence
  prev_ci <- prop_with_ci(TP + FN, N) # sample prevalence CI

  # Additional metrics
  add_list <- list()
  if (add_additional) {
    # Likelihood Ratios (CIs via log method)
    # LR+ = sens / (1 - spec)
    # LR- = (1 - sens) / spec
    # Var(log(LR+)) = Var(log(sens)) + Var(log(1-spec))
    # Approx Var(p) ~ p(1-p)/n; Delta method.
    sens_p <- sens$estimate; spec_p <- spec$estimate
    sens_n <- TP + FN; spec_n <- TN + FP

    lr_plus <- if (is.na(sens_p) || is.na(spec_p) || spec_p == 1) NA else sens_p / (1 - spec_p)
    lr_minus <- if (is.na(sens_p) || is.na(spec_p) || spec_p == 0) NA else (1 - sens_p) / spec_p

    lr_plus_ci <- c(NA, NA)
    lr_minus_ci <- c(NA, NA)

    z <- qnorm(1 - (1-ci_level)/2)

    if (!is.na(lr_plus) && !is.na(sens_p) && !is.na(spec_p) &&
        sens_p > 0 && sens_p < 1 && spec_p > 0 && spec_p < 1 &&
        sens_n > 0 && spec_n > 0 && (1 - spec_p) > 0) {
      var_log_sens <- (1 - sens_p)/(sens_p * sens_n)
      var_log_1minus_spec <- spec_p / ((1 - spec_p) * spec_n)
      se_log_lr_plus <- sqrt(var_log_sens + var_log_1minus_spec)
      lr_plus_ci <- exp(log(lr_plus) + c(-1,1)*z*se_log_lr_plus)
    }

    if (!is.na(lr_minus) && !is.na(sens_p) && !is.na(spec_p) &&
        sens_p > 0 && sens_p < 1 && spec_p > 0 && spec_p < 1 &&
        sens_n > 0 && spec_n > 0) {
      var_log_1minus_sens <- sens_p / ((1 - sens_p) * sens_n)
      var_log_spec <- (1 - spec_p)/(spec_p * spec_n)
      se_log_lr_minus <- sqrt(var_log_1minus_sens + var_log_spec)
      lr_minus_ci <- exp(log(lr_minus) + c(-1,1)*z*se_log_lr_minus)
    }

    # F1 score (no simple exact CI; use Delta approx on logit? We'll skip CI or bootstrap; provide NA or Wald)
    precision_p <- ppv$estimate
    recall_p <- sens_p
    f1 <- if (is.na(precision_p) || is.na(recall_p) || (precision_p + recall_p) == 0) NA else
      2 * precision_p * recall_p / (precision_p + recall_p)
    # Provide naive bootstrap possibility? Keep simple: no CI
    f1_ci <- c(NA, NA)

    add_list <- list(
      lr_positive = lr_plus,
      lr_positive_lower_ci = lr_plus_ci[1],
      lr_positive_upper_ci = lr_plus_ci[2],
      lr_negative = lr_minus,
      lr_negative_lower_ci = lr_minus_ci[1],
      lr_negative_upper_ci = lr_minus_ci[2],
      f1 = f1,
      f1_lower_ci = f1_ci[1],
      f1_upper_ci = f1_ci[2]
    )
  }

  # "Rule-out" custom metrics the original code had:
  # Original code used confusion_mtx[1] and confusion_mtx[3] which are column-major indices.
  # We'll interpret "Correct rule-out (ERCP avoided)" as TN and "False rule-out (IDCs missed)" as FN.
  correct_rule_out <- TN
  false_rule_out   <- FN
  correct_rule_out_prop <- if (N == 0) NA else correct_rule_out / N
  false_rule_out_prop   <- if (N == 0) NA else false_rule_out / N

  correct_rule_out_ci <- prop_with_ci(correct_rule_out, N)
  false_rule_out_ci   <- prop_with_ci(false_rule_out, N)

  # Assemble result
  res_list <- c(
    # Counts
    TP = TP, FP = FP, FN = FN, TN = TN, N = N,
    # Prevalence
    prevalence_sample = prev_sample,
    prevalence_sample_lower_ci = prev_ci$lower,
    prevalence_sample_upper_ci = prev_ci$upper,
    prevalence_used = prev,
    # Sensitivity
    sensitivity = sens$estimate,
    sensitivity_lower_ci = sens$lower,
    sensitivity_upper_ci = sens$upper,
    # Specificity
    specificity = spec$estimate,
    specificity_lower_ci = spec$lower,
    specificity_upper_ci = spec$upper,
    # PPV
    ppv = ppv$estimate,
    ppv_lower_ci = ppv$lower,
    ppv_upper_ci = ppv$upper,
    # NPV
    npv = npv$estimate,
    npv_lower_ci = npv$lower,
    npv_upper_ci = npv$upper,
    # Accuracy
    accuracy = acc$estimate,
    accuracy_lower_ci = acc$lower,
    accuracy_upper_ci = acc$upper,
    # Rule-out metrics
    `Correct rule-out (ERCP avoided) [n]` = correct_rule_out,
    `Correct rule-out (ERCP avoided) [%]` = correct_rule_out_prop * 100,
    `Correct rule-out (ERCP avoided) [%]_lower_ci` = correct_rule_out_ci$lower * 100,
    `Correct rule-out (ERCP avoided) [%]_upper_ci` = correct_rule_out_ci$upper * 100,
    `False rule-out (IDCs missed) [n]` = false_rule_out,
    `False rule-out (IDCs missed) [%]` = false_rule_out_prop * 100,
    `False rule-out (IDCs missed) [%]_lower_ci` = false_rule_out_ci$lower * 100,
    `False rule-out (IDCs missed) [%]_upper_ci` = false_rule_out_ci$upper * 100
  )

  if (add_additional) {
    res_list <- c(res_list, add_list)
  }

  # Rounding (keep counts & LRs un-rounded or lightly rounded)
  round_if_prob <- function(name, value) {
    if (is.na(value)) return(NA_real_)
    # Do not round counts (integers)
    if (grepl("\\[n\\]$", name) || name %in% c("TP","FP","FN","TN","N")) {
      return(value)
    }
    # Likelihood ratios: 3 decimals
    if (grepl("^lr_", name)) {
      return(round(value, 3))
    }
    # Proportions (0-1)
    if (value >= 0 && value <= 1 && !grepl("\\[%\\]", name)) {
      return(round(value, 3))
    }
    # Percentages
    if (grepl("\\[%\\]", name)) {
      return(round(value, 2))
    }
    # F1 etc (0-1)
    if (name == "f1" || grepl("^f1_", name)) {
      return(round(value, 3))
    }
    return(value)
  }

  res_list <- mapply(round_if_prob, names(res_list), res_list, SIMPLIFY = TRUE, USE.NAMES = TRUE)

  res_list
}




# Summarize and add lab data 
add_lab = function (x_id) {
    # This function selects lab data within 24h before CTPA,
    # summarizes and adds it to the clinical data
    x_tab = visit_table_v0 %>%
      filter(pid == x_id)
    x_timing_cta = x_tab$timing_cta
    x_lab_data = aggregated_lab_data %>%
      filter(id == x_id & ((x_timing_cta - .$time) %in% c(0:1)))
    is_empty = FALSE
    if(nrow(x_lab_data) == 0) {
      x_lab_data = aggregated_lab_data[1,]
      for (i in 1:ncol(x_lab_data))
      {
        x_lab_data[,i] = -1
        x_lab_data[1,3] = NA}
      x_lab_data$id = x_id
      is_empty = TRUE
    }
    x_summary = x_lab_data %>%
      select(-c(1,2)) %>%
      summary() %>%
      apply(c(1,2),
            function(x){suppressWarnings(as.numeric(sub('.*:', '', x)))})
    rownames(x_summary) = c("min",
                            "q1",
                            "median",
                            "mean",
                            "q3",
                            "max",
                            "NA")
    x_summary = x_summary %>%
      as_tibble(rownames = "rnms") %>%
      pivot_longer(cols = 2:ncol(.)) %>%
      add_column(val_name = paste("lab", .$rnms, .$name, sep = "_"),
                 .before = 1) %>%
      select("val_name", "value") %>%
      mutate(val_name = gsub(pattern = " ", replacement = "", val_name))
    x_list = as.list(x_summary$value)
    names(x_list) = x_summary$val_name
    if(is_empty == TRUE){x_list = lapply(x_list, function(x){NA})}
    x_list = lapply(x_list, function(x){ifelse(x>=0, x, NA)}) # remove neg. values
    x_list = x_list %>% data.frame() %>% as_tibble()
    x_tab = x_tab %>% add_column(x_list)
    x_tab
}


# Get ROC dataframe
make_roc_df = function(data, score, score_name){
    roc_empirical = rocit(score = score,
                          class = data$has_idc,
                          negref = "no")
    ci95=ciAUC(roc_empirical)
    roc_df = data_frame(TPR = roc_empirical$TPR, 
                        FPR = roc_empirical$FPR, 
                        Score = score_name, 
                        AUC = round(roc_empirical$AUC, digits = 3),
                        "CI95_upper" = round(ci95$lower, digits = 3),
                        "CI95_lower" = round(ci95$upper, digits = 3))
    return(roc_df)
}
```




## Import data
Import data and perform some pre-processing (filter deleted database rows, define column data types, etc.)
```{r}
# Import data
all_data = read_csv("data/input_data_tab1.csv")
test_df = read_csv("data/test_df_pandas.csv")
model_cols_df = read_csv("data/model_cols_df.csv")
```

Add columns
```{r}
all_data$has_mrcp = !is.na(all_data$mrcp_date)
all_data$has_eus = !is.na(all_data$eus_date)
all_data$has_ercp = !is.na(all_data$ercp_date)
all_data$has_abdominal_sonography = !is.na(all_data$abdominal_sonography_date)
all_data$has_ct_mri = !is.na(all_data$ct_mri_date)
```


# Results

## Patient characteristics

### Tabel 1 Patient characteristics
General characteristics of the cohort.
```{r include=FALSE}
colnames(all_data) = make.names(colnames(all_data))
colnames(all_data)
```

```{r}
all_data$clinical_data_pain_onset_in_days_before_admission = as.numeric(all_data$clinical_data_pain_onset_in_days_before_admission)
all_data$eus_ercp_mrcp_or_as_dhc_width_in_mm = as.numeric(all_data$eus_ercp_mrcp_or_as_dhc_width_in_mm)


summary_data = all_data[, -c(1:2)] %>% select(starts_with(c("has_idc",
                                                 "baseline_",
                                                 "lab_values_admission_",
                                                 "clinical_data_",
                                                 "eus_ercp_mrcp_or_as_dhc_width_in_mm",
                                                 "has_")) & !contains("date"))
summary_table = gtsummary::tbl_summary(summary_data, by = has_idc, type = list(eus_ercp_mrcp_or_as_dhc_width_in_mm ~ "continuous", clinical_data_pain_onset_in_days_before_admission ~ "continuous")) 
summary_table = gtsummary::add_p(summary_table) 
# summary_table = gtsummary::add_ci(summary_table)

summary_table
```



## Prediction performances

### TFDF model parameters
TensorFlow Decision Forests (TFDF) for the detection of IDC.

```{r}
model_cols_df
```

### Figure 2
Variable importances
```{r}
feature_importances = read_csv("data/feature_importances.csv")
feature_importances$feature_names[feature_importances$feature_names == "alkaline_phosphatase_in_u_l"] = "alkaline phosphatase"
feature_importances$feature_names[feature_importances$feature_names == "got_ast_in_u_l"] = "got/ast"
feature_importances$feature_names[feature_importances$feature_names == "nicotine_use"] = "nicotine use"

feature_importances$feature_names = str_to_upper(feature_importances$feature_names)
feature_importances$feature_names = sapply(feature_importances$feature_names, function(x) strsplit(x, "_")[[1]][[1]])

# Rename some features:
feature_importances$feature_names[feature_importances$feature_names == "GPT"] = "ALT"
feature_importances$feature_names[feature_importances$feature_names == "GOT/AST"] = "AST"

# Plot
p = ggplot2::ggplot(feature_importances, aes(y = feature_importances, x = reorder(feature_names, feature_importances), color = str_to_title(feature_type))) +
    geom_point() +
    geom_segment(aes(x=feature_names,xend=feature_names,y=0,yend=feature_importances)) +
    labs(# title = "Variable importance",
         y = "Importance [inverted mean minimal depth]",
         x = "Variable",
         color = "Type") +
    coord_flip() + 
    theme_minimal() +
    theme(legend.position="right")
  
ggplot2::ggsave("Figure 2. Variable importance.pdf", p, device = "pdf", height = 8, width = 10)
p
```

**Figure 2. Variable importance.** The importance of each variable included in the TFDF-model as estimated by the inverted mean minimal depth of each variable in the decision tree.

```{r}
# Set prevalence of idc
prevalence_idc = sum(all_data$has_idc == "1")/nrow(all_data)
```



### TFDF-model

Tensorflow random forests model.
```{r}
test_df$has_idc = as.character(test_df$has_idc)
test_df$has_idc[test_df$has_idc == "1"] = "yes"
test_df$has_idc[test_df$has_idc == "0"] = "no"
test_df$pred_idc = as.character(test_df$pred_idc)
test_df$pred_idc[test_df$pred_idc == TRUE] = "yes"
test_df$pred_idc[test_df$pred_idc == FALSE] = "no"

table_tfdf = table(test_df[, c("pred_idc", "has_idc")])
confusionMatrix(table_tfdf, prevalence = prevalence_idc, positive = "yes")
```



### CBD model
CBD 8 mm or larger
```{r}
test_df$cbd_gte_6mm = "no"
test_df$cbd_gte_6mm[test_df$dhc_in_mm >= 6] = "yes"
test_df$cbd_gte_7mm = "no"
test_df$cbd_gte_7mm[test_df$dhc_in_mm >= 7] = "yes"
test_df$cbd_gte_8mm = "no"
test_df$cbd_gte_8mm[test_df$dhc_in_mm >= 8] = "yes"

# table_dhc = table(test_df[, c("cbd_gte_6mm", "has_idc")])
# confusionMatrix(table_dhc, prevalence = prevalence_idc, positive = "yes")
# table_dhc = table(test_df[, c("cbd_gte_7mm", "has_idc")])
# confusionMatrix(table_dhc, prevalence = prevalence_idc, positive = "yes")
table_dhc = table(test_df[, c("cbd_gte_8mm", "has_idc")])
confusionMatrix(table_dhc, prevalence = prevalence_idc, positive = "yes")


test_df$bili_gte_4 = "no"
test_df$bili_gte_4[test_df$lab_values_admission_bilirubin_in_mg_dl >= 4] = "yes"

table_bil = table(test_df[, c("bili_gte_4", "has_idc")])


test_df$cbd_gte_8mm_AND_bili_gte_4 = "no"
test_df$cbd_gte_8mm_AND_bili_gte_4[test_df$lab_values_admission_bilirubin_in_mg_dl >= 4 & test_df$dhc_in_mm >= 8] = "yes"

table_dhc_and_bil = table(test_df[, c("cbd_gte_8mm_AND_bili_gte_4", "has_idc")])



```

<!-- ### Neuronal networks -->
<!-- Confusion matrix of the neuronal network model. Of note, this only comprises data from the test dataset. -->
<!-- ```{r} -->
<!-- mtx = matrix(c(5,3,3,61), nrow = 2) -->
<!-- dimnames(mtx) = list(c("FALSE", "TRUE"), c("FALSE", "TRUE")) -->
<!-- table_tfnn = as.table(mtx) -->
<!-- confusionMatrix(table_tfnn, positive = pos, prevalence = 0.147) -->
<!-- ``` -->

### Subgroups
```{r}
stones = filter(test_df, (clinical_data_stone_microlithiasis_sludge %in% c(1,4,5,7)))
sludge_mic = filter(test_df, (clinical_data_stone_microlithiasis_sludge %in% c(2,3,6)))
```


#### Stones
##### TFDF-model

Tensorflow random forests model.
```{r}
stones$has_idc = as.character(stones$has_idc)
stones$has_idc[stones$has_idc == "1"] = "yes"
stones$has_idc[stones$has_idc == "0"] = "no"
stones$pred_idc = as.character(stones$pred_idc)
stones$pred_idc[stones$pred_idc == TRUE] = "yes"
stones$pred_idc[stones$pred_idc == FALSE] = "no"

table_tfdf_stones = table(stones[, c("pred_idc", "has_idc")])
confusionMatrix(table_tfdf_stones, prevalence = prevalence_idc, positive = "yes")
```


##### CBD model
CBD 8 mm or larger
```{r}
stones$cbd_gte_8mm = "no"
stones$cbd_gte_8mm[stones$dhc_in_mm >= 8] = "yes"

table_dhc_stones = table(stones[, c("cbd_gte_8mm", "has_idc")])
confusionMatrix(table_dhc_stones, prevalence = prevalence_idc, positive = "yes")
```


#### Sludge/microlithiasis
##### TFDF-model

Tensorflow random forests model.
```{r}
sludge_mic$has_idc = as.character(sludge_mic$has_idc)
sludge_mic$has_idc[sludge_mic$has_idc == "1"] = "yes"
sludge_mic$has_idc[sludge_mic$has_idc == "0"] = "no"
sludge_mic$pred_idc = as.character(sludge_mic$pred_idc)
sludge_mic$pred_idc[sludge_mic$pred_idc == TRUE] = "yes"
sludge_mic$pred_idc[sludge_mic$pred_idc == FALSE] = "no"

table_tfdf_sludge_mic = table(sludge_mic[, c("pred_idc", "has_idc")])
confusionMatrix(table_tfdf_sludge_mic, prevalence = prevalence_idc, positive = "yes")
```


##### CBD model
CBD 8 mm or larger
```{r}
sludge_mic$cbd_gte_8mm = "no"
sludge_mic$cbd_gte_8mm[sludge_mic$dhc_in_mm >= 8] = "yes"

table_dhc_sludge_mic = table(sludge_mic[, c("cbd_gte_8mm", "has_idc")])
confusionMatrix(table_dhc_sludge_mic, prevalence = prevalence_idc, positive = "yes")
```


### Summary table
```{r}
perf_summary = sapply(list(
            "TFDF-Model" = table_tfdf,
            "CBD & BILIRUBIN" = table_dhc_and_bil,
            "CBD" = table_dhc,
            "BILIRUBIN" = table_bil,
            "Subgroup Stones: TFDF-Model" = table_tfdf_stones,
            "Subgroup Stones: CBD" = table_dhc_stones,
            "Subgroup Sludge/microlithiasis: TFDF-Model" = table_tfdf_sludge_mic,
            "Subgroup Sludge/microlithiasis: CBD" = table_dhc_sludge_mic),
       get_perf, 
       prevalence = prevalence_idc, 
       simplify = TRUE, USE.NAMES = TRUE)


perf_summary %>%
  data.frame() %>%
  kbl() %>%
  kable_styling()

```



### Figure 3
```{r fig.height=8, fig.width=8, dpi=600}


# Generate mosaic data
mosaic_data <- test_df %>%
    pivot_longer(c(pred_idc, cbd_gte_8mm_AND_bili_gte_4)) %>%
    filter(!is.na(value)) %>%
    mutate(name = gsub("-", " ", name)) %>%
    mutate(name = gsub("_", " ", name)) %>%
    mutate(name = str_to_upper(name)) %>%
    add_column(rule_in_out = ifelse(.$value == "no" & .$has_idc == "no",
                                    "Correct rule-out",
                             ifelse(.$value == "yes" & .$has_idc == "yes",
                                    "Correct rule-in",
                             ifelse(.$value == "no" & .$has_idc == "yes",
                                    "False rule-out",
                             ifelse(.$value == "yes" & .$has_idc == "no",
                                    "False rule-in", NA))))) %>%
    mutate(rule_in_out = as.factor(rule_in_out)) %>%
    mutate(has_idc = as.factor(has_idc)) %>%
    add_column(score_rule = sapply(as.logical(.$value == "yes"),
                                   function(x) ifelse(x, "rule-in", "rule-out"))) 

fill_pal <- pal_jco("default")(4)
pA <- ggplot(data = mosaic_data) +
    geom_mosaic(aes(x = product(score_rule, has_idc)),
                fill = rep(fill_pal, 2),
                offset = 0.02) +
    labs(x = "IDC diagnosed by ERC/MRC/EUS/AS",
         y = "Score result",
         fill = "Rule-out by score") +
    theme_minimal() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()) +
    facet_wrap("name")

# Add custom legend
legend <- ggplot(mosaic_data, aes(x = rule_in_out, fill = rule_in_out)) + 
    geom_bar() +
    scale_fill_manual(values = fill_pal[c(3, 2, 1, 4)]) +
    labs(fill = "")
legend <- get_legend(legend)
legend <- as_ggplot(legend)
pA <- ggarrange(pA, legend, widths = c(5,1))

roc_df <- rbind(make_roc_df(test_df, test_df$pred_prob, "PRED IDC"),
                make_roc_df(test_df, test_df$lab_values_admission_bilirubin_in_mg_dl, "BILIRUBIN"),
                make_roc_df(test_df, test_df$dhc_in_mm, "CBD DIAMETER"))

plt_df <- roc_df %>%
    filter(Score != "PRED IDC") %>%
    mutate(Score = as.factor(as.character(Score))) %>%
    group_by(Score) %>%
    summarize(AUC = first(AUC),
              CI95_upper = first(CI95_upper),
              CI95_lower = first(CI95_lower))

# Offsets for vertical stacking
y_offsets <- seq(0.1, 0.05, length.out = nrow(plt_df))

pB <- ggplot(filter(roc_df, Score != "PRED IDC"), aes(x = FPR, y = TPR, color = Score)) +
    geom_line() +
    geom_line(data = data.frame(x = c(0, 1), y = c(0, 1)),
              mapping = aes(x = x, y = y, colour = NULL),
              lty = "dashed") +
    labs(x = "1 - specificity (FPR)",
         y = "Sensitivity (TPR)") +
    scale_color_jco() +
    theme_minimal() +
    theme(legend.position = "bottom") +
    geom_text(data = plt_df,
              aes(x = 0.3, y = y_offsets,
                  label = paste0("AUC: ", AUC, " (CI 95%: ", CI95_upper, ", ", CI95_lower, ")")),
              hjust = 0, vjust = 0, size = 3.5)

plt_df2 <- roc_df %>%
    filter(Score == "PRED IDC") %>%
    mutate(Score = as.factor(as.character(Score)))

pC <- ggplot(filter(plt_df2, Score == "PRED IDC"), aes(x = FPR, y = TPR, color = Score)) +
    geom_line() +
    geom_line(data = data.frame(x = c(0, 1), y = c(0, 1)),
              mapping = aes(x = x, y = y, colour = NULL),
              lty = "dashed") +
    labs(x = "1 - specificity (FPR)",
         y = "Sensitivity (TPR)") +
    scale_color_manual(values = pal_jco("default")(1)) +
    theme_minimal() +
    theme(legend.position = "bottom") +
    geom_text(data = plt_df2,
              aes(x = 0.3, y = 0.1,
                  label = paste0("AUC: ", AUC, " (CI 95%: ", CI95_upper, ", ", CI95_lower, ")")),
              hjust = 0, vjust = 0, size = 3.5)

plot_data <- all_data
plot_data$has_idc <- as.character(plot_data$has_idc)
plot_data$has_idc[plot_data$has_idc == "1"] <- "yes"
plot_data$has_idc[plot_data$has_idc == "0"] <- "no"

pBC <- ggarrange(pB, pC, labels = c("B", "C"), nrow = 1)
p <- ggarrange(pA, pBC, labels = c("A", ""), nrow = 2, heights = c(5, 6))
ggplot2::ggsave("Figure 3. Comparison of decision rules.pdf", p, device = "pdf", height = 8, width = 10)
# Print the final plot
print(p)
```

**Figure 3. Comparison of decision rules.** (A) Mosaic diagram with each field’s size corresponding to the count of correct and false rule-in and -out (i.e. to the four fields of the respective contingency tables). False rule-out in red signifies missed IDCs and the correct rule-out in dark-blue corresponds to correctly avoided ERCs. (B) Two ROC curves were plotted for the TFDF model and the CBD diameter.

### Figure 4
Model performance for subgroups
```{r fig.height=8, fig.width=10, dpi=600}
roc_df <- rbind(make_roc_df(stones, stones$pred_prob, "PRED IDC"),
                make_roc_df(stones, stones$lab_values_admission_bilirubin_in_mg_dl, "BILIRUBIN"),
                make_roc_df(stones, stones$dhc_in_mm, "CBD DIAMETER"))


pA = ggplot(roc_df, aes(x = FPR, y = TPR)) +
    geom_line() +
    geom_line(data = data.frame(x = c(0,1), y = c(0,1)),
              mapping = aes(x=x, y=y, colour = NULL),
              lty = "dashed") +
    labs(x = "1 - specificity (FPR)",
         y = "Gallstones; Sensitivity (TPR)") +
    theme_minimal() +
    facet_wrap("Score") + 
    geom_text(data = roc_df,
              aes(1, 1, 
                  label = paste0("AUC: ", AUC, " (CI 95%: ", CI95_upper, ", ", CI95_lower, ")")),
              hjust = 1.6,
              vjust = 2,
              size=3) 

roc_df = rbind(
                make_roc_df(sludge_mic, sludge_mic$pred_prob, "PRED IDC"),
                make_roc_df(sludge_mic, sludge_mic$lab_values_admission_bilirubin_in_mg_dl, "BILIRUBIN"),
                make_roc_df(sludge_mic, sludge_mic$dhc_in_mm, "CBD DIAMETER")
              )


pB = ggplot(roc_df, aes(x = FPR, y = TPR)) +
    geom_line() +
    geom_line(data = data.frame(x = c(0,1), y = c(0,1)),
              mapping = aes(x=x, y=y, colour = NULL),
              lty = "dashed") +
    labs(x = "1 - specificity (FPR)",
         y = "Sludge/Microlithiasis; Sensitivity (TPR)") +
    theme_minimal() +
    facet_wrap("Score") + 
    geom_text(data = roc_df,
              aes(1, 1, 
                  label = paste0("AUC: ", AUC, " (CI 95%: ", CI95_upper, ", ", CI95_lower, ")")),
              hjust = 1.6,
              vjust = 2,
              size=3) 


plot_data = all_data
plot_data$has_idc = as.character(plot_data$has_idc)
plot_data$has_idc[plot_data$has_idc == "1"] = "yes"
plot_data$has_idc[plot_data$has_idc == "0"] = "no"

p = ggarrange(pA, pB, labels = c("A", "B"), nrow = 2, heights = c(1,1))
ggplot2::ggsave("Figure 4. Comparison of decision rules for microlithiasis and sludge.pdf", p, device = "pdf", height = 8, width = 10)
p
```
**Figure 4. Comparison of decision rules for microlithiasis and sludge.** Model performances were computed for two subgroups: patients with gallstones (A, n=`r nrow(stones)`) and with sludge and/or microlithiasis (B, n=`r nrow(sludge_mic)`). For each subgroup, two ROC curves were plotted for the TFDF model and the CBD diameter.


### Figure 5?
Something like that?

```{r fig.height=4, fig.width=12, dpi=600}

p2 = ggplot(plot_data, aes(x=has_idc, y = eus_ercp_mrcp_or_as_dhc_width_in_mm,
                               fill = has_idc)) +
    geom_violin(width = 0.8, alpha = 0.5) +
    geom_boxplot(width = 0.1, alpha = 0.5) +
    geom_hline(yintercept=11, linetype="dashed") +
    geom_hline(yintercept=8, linetype="dotted") +
    theme_minimal() +
    theme(legend.position = "none") +
    labs(# title = "D-dimer",
         y = "CBD [mm]",
         x = "Diagnosis of IDC [yes/no]",
         #fill = "PE"
         ) +
    scale_y_log10() +
    geom_signif(
        comparisons = list(c("yes", "no")),
        map_signif_level = FALSE
        ) 

p5 = ggscatter(plot_data,
               x = "lab_values_admission_bilirubin_in_mg_dl",
               y = "eus_ercp_mrcp_or_as_dhc_width_in_mm",
          add = "reg.line",color = "has_idc", shape = "has_idc",palette = "jco",
          add.params = list(color = "blue", fill = "lightgray"),
          conf.int = TRUE) +
    scale_x_log10() +
    scale_y_log10() +
    stat_cor(method = "spearman", label.x = 0.7, label.y = 1.5) +
    labs(x = "Bilirubin [mg/dl]",
         y = "CBD [mm]",
         color = "IDC",
         shape = "IDC"
    ) +
    geom_hline(yintercept=11, linetype="dashed") +
    geom_hline(yintercept=8, linetype="dotted") +
    geom_vline(xintercept=4, linetype="dashed") +
    geom_vline(xintercept=1, linetype="dotted") +
    theme_minimal() +
    theme(legend.position = "right")

p6 = ggplot(plot_data, aes(x=has_idc, y = lab_values_admission_bilirubin_in_mg_dl,
                               fill = has_idc)) +
    geom_violin(width = 0.8, alpha = 0.5) +
    geom_boxplot(width = 0.1, alpha = 0.5) +
    geom_hline(yintercept=4, linetype="dashed") +
    geom_hline(yintercept=1, linetype="dotted") +
    theme_minimal() +
    theme(legend.position = "none") +
    labs(
         y = "Bilirubin [mg/dl]",
         x = "Diagnosis of IDC [yes/no]",
         ) +
    scale_y_log10() +
    geom_signif(
        comparisons = list(c("yes", "no")),
        map_signif_level = FALSE
        ) 

pA = ggarrange(set_palette(p2, "jco"),
               set_palette(p6, "jco"),
               set_palette(p5, "jco"), labels = c("A", "B", "C"),
               ncol = 3,
               widths = c(3,3,8))

ggplot2::ggsave("Figure 5. CBD diameter.pdf", pA, device = "pdf", height = 5, width = 12)
pA
```

**Figure 5. CBD diameter.** (A) Violin plot of CBD diameters comparing IDC versus no IDC. The p-value is calculated using a two-sided Mann-Whitney test. The dashed line corresponds to a CBD of 11 mm; the dotted line to 7 mm. (B) Spearman correlation of CBD diameter and gGT. The vertical dashed line corresponds to a gGT of 50 U/l; values < 1.5 pg/ml are considered normal. Note the log-transformed axes for gGT in (B).

# Appendix

## TFDF Model partial tree
first three levels
````{=html}
```{r, echo=FALSE, results='asis'}
xfun::file_string('data/tree-levels-1-3.svg')
```
````

## TFDF Model full tree
````{=html}
```{r, echo=FALSE, results='asis'}
xfun::file_string('data/tree-levels-all.svg')
```
````



